---
title: "Homework 3"
author: "Saryu Patel"
date: "10/15/2021"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
theme_set(theme_minimal())

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r}
library(p8105.datasets)
data("instacart")
```

The `instacart` dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Some key variables are `product_name`, `aisle`, and `department`. Below is a table summarizing the top 10 most frequently bought items in the dataset. 

```{r echo=FALSE}
instacart %>% 
  group_by(product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  head(10) %>% 
  knitr::kable()
```

Bananas are the most frequently bought item.

The most popular department to shop from is the `r instacart %>% group_by(department) %>% summarize(n_obs = n()) %>% arrange(desc(n_obs)) %>% head(1) %>% select(department)` department.

There are `r nrow(instacart %>% select(aisle) %>% unique())` aisles. Below is a table summarizing the top 10 aisles where the most items are ordered from.

```{r echo=FALSE}
pop_aisles <- 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) 
pop_aisles %>% 
  arrange(desc(n_obs)) %>% 
  head(10) %>% 
  knitr::kable()
```

The fresh vegetables and fresh fruits aisles are the most common aisles from where items are ordered. 

Below is a plot that shows the number of items ordered in each aisle, limited to aisles with more than 10000 items ordered. 

```{r echo=FALSE}
pop_aisles %>% 
  filter(n_obs > 10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_obs, .desc = TRUE)) %>% 
  ggplot(aes(x = aisle, y = n_obs)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Number of Items Ordered in the Most Popular Aisles",
       x = "Aisle",
       y = "Number of Items") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

It appears that the fresh vegetables and fresh fruits aisles are by far the most popular.

Below is a table showing the three most popular items in the isles "baking ingredients", "dog food care", and "packaged vegetables fruits".

```{r echo=FALSE, message=FALSE}
instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  slice(1:3) %>% 
  knitr::kable()
```

Light Brown Sugar is the most popular item from the baking ingredients aisle, Organix Dog Food is the most popular item from the dog food care aisle, and Organic Baby Spinach is the most popular item from the packaged vegetables fruits aisle.

Below is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream were ordered on each day of the week.

```{r echo=FALSE, warning=FALSE, message=FALSE}
instacart %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(names_from = order_dow,
              values_from = mean_hour) %>% 
  knitr::kable()
```

Apart from day of the week number 5, Pink Lady Apples tend to be bought at a mean time that is earlier than the mean time at which Coffee Ice Creams.

### Problem 2

```{r}
data("brfss_smart2010")
```

```{r warning=FALSE}
brfss_df <- 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health", !response %in% c("Don't know/Not Sure", "Refused")) %>% 
  mutate(response = forcats::fct_relevel(response, "Poor", "Fair", "Good", "Very good", "Excellent"))
```

In 2002, the following states were observed at 7 or more locations:

```{r echo=FALSE}
brfss_df %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarize(n_locations = n_distinct(locationdesc)) %>% 
  filter(n_locations >= 7) %>% 
  select(locationabbr) %>% 
  knitr::kable()
```

In 2010, the following states were observed at 7 or more locations:

```{r}
brfss_df %>% 
  filter(year == 2010) %>% 
  group_by(locationabbr) %>% 
  summarize(n_locations = n_distinct(locationdesc)) %>% 
  filter(n_locations >= 7) %>% 
  select(locationabbr) %>% 
  knitr::kable()
```

The number of states that were observed at 7 or more locations increased from 2007 to 2010. CT is the only state that was observed at 7 or more locations in 2007 but not in 2010. 

```{r message=FALSE}
excellent_df <- 
  brfss_df %>% 
  filter(response == "Excellent") %>% 
  select(year, locationabbr, data_value) %>% 
  group_by(locationabbr, year) %>% 
  summarize(avg_value = mean(data_value))
```

Below is a plot showing the average value over time within a state.

```{r echo=FALSE, warning=FALSE}
excellent_df %>% 
  ggplot(aes(x = year, y = avg_value, group = locationabbr, color = locationabbr)) +
  geom_line() +
  labs(title = "Average Data Value Over Time by State",
       x = "Year",
       y = "Average Value",
       color = "State")
```

Below is a two-panel plot showing the distribution of `data_value` for responses among locations in NY State for the years 2006 and 2010.

```{r echo=FALSE, warning=FALSE}
brfss_df %>% 
  filter(locationabbr == "NY",
         year %in% c(2006, 2010)) %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ year)
```

There appears to be a lower average value for the response "Excellent" in 2006 than there is in 2010. There is a higher density of "Poor" values in 2006 than there is in 2010. 

### Problem 3

```{r message=FALSE, warning=FALSE}
accel_df <- 
  read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(activity_1:activity_1440, 
               names_to = "min_of_day",
               values_to = "activity_count") %>% 
  mutate(min_of_day = str_sub(min_of_day, 10),
         min_of_day = as.integer(min_of_day),
         day = forcats::fct_relevel(day, "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))

i = 1
for (row in accel_df$day) {
  if (row == "Saturday" || row == "Sunday") {
    accel_df$type_of_day[i] = "weekend"
    i = i + 1
  } else {
    accel_df$type_of_day[i] = "weekday"
    i = i + 1
  }
}
```

The resulting dataset has `r nrow(accel_df)` observations and `r ncol(accel_df)` variables. The most important variable is `activity_count`, and the overall mean of this variable is `r round(mean(accel_df$activity_count), 2)`. 

Below is a table summarizing the total activity over each day. 

```{r echo=FALSE, message=FALSE}
accel_df %>% 
  group_by(day_id, day) %>% 
  summarize(total_activity = sum(activity_count)) %>% 
  knitr::kable()
```

Fridays, Saturdays, and Sundays tend to have a greater total activity. There are two Saturdays where the total activity is only the total number of minutes in a day. 

```{r}
accel_df %>% 
  ggplot(aes(x = min_of_day, y = activity_count, group = day_id, color = day)) +
  geom_line()
```

The highest activity counts occur in the evenings, regardless of day. There is some spike in activity in the afternoon on Saturdays and Sundays. Fridays and Sundays have a spike of activity in the morning, with Sunday activity being later in the morning.