---
title: "Homework 3"
author: "Saryu Patel"
date: "10/15/2021"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
```

### Problem 1

```{r}
library(p8105.datasets)
data("instacart")
```

The `instacart` dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Some key variables are `product_name`, `aisle`, and `department`. Below is a table summarizing the top 10 most frequently bought items in the dataset. 

```{r echo=FALSE}
instacart %>% 
  group_by(product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  head(10) %>% 
  knitr::kable()
```

Bananas are the most frequently bought item.

The most popular department to shop from is the `r instacart %>% group_by(department) %>% summarize(n_obs = n()) %>% arrange(desc(n_obs)) %>% head(1) %>% select(department)` department.

There are `r nrow(instacart %>% select(aisle) %>% unique())` aisles. Below is a table summarizing the top 10 aisles where the most items are ordered from.

```{r echo=FALSE}
pop_aisles <- 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) 
pop_aisles %>% 
  arrange(desc(n_obs)) %>% 
  head(10) %>% 
  knitr::kable()
```

The fresh vegetables and fresh fruits aisles are the most common aisles from where items are ordered. 

Below is a plot that shows the number of items ordered in each aisle, limited to aisles with more than 10000 items ordered. 

```{r echo=FALSE}
pop_aisles %>% 
  filter(n_obs > 10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_obs, .desc = TRUE)) %>% 
  ggplot(aes(x = aisle, y = n_obs)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Number of Items Ordered in the Most Popular Aisles",
       x = "Aisle",
       y = "Number of Items") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

It appears that the fresh vegetables and fresh fruits aisles are by far the most popular.

Below is a table showing the three most popular items in the isles "baking ingredients", "dog food care", and "packaged vegetables fruits".

```{r echo=FALSE, message=FALSE}
instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs)) %>% 
  slice(1:3) %>% 
  knitr::kable()
```

Light Brown Sugar is the most popular item from the baking ingredients aisle, Organix Dog Food is the most popular item from the dog food care aisle, and Organic Baby Spinach is the most popular item from the packaged vegetables fruits aisle.

Below is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream were ordered on each day of the week.

```{r echo=FALSE, warning=FALSE, message=FALSE}
instacart %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(names_from = order_dow,
              values_from = mean_hour) %>% 
  knitr::kable()
```

Apart from day of the week number 5, Pink Lady Apples tend to be bought at a mean time that is earlier than the mean time at which Coffee Ice Creams.

### Problem 2


### Problem 3